Manuscript

### 2025-09-14 Manuscript_MD_ScoreReLiability.qmd
update manuscript up to methods

---
title: "Manuscript"
bibliography: ReLiability_references.bib
csl: styles/apa-6th-edition.csl
format: html
editor: 
  markdown: 
    wrap: 72
---

@@

# relevant libraries required for this script
packages <- c("ggplot2", "here", "metafor", "gridExtra", "knitr", "magrittr", "dplyr")

@@

@@

Similarly, it has been argued that heterogeneity in effect sizes is an
indicator of the theory’s “completeness” surrounding the phenomenon.
Linden and Hönekopp argue that “low (as opposed to high) heterogeneity
reflects a more advanced understanding of the subject matter being
studied” [-@LindenHönekopp, p. 2]. Similarly, von Hippel and Schuetze
[-@vonHippel2023] argue that heterogeneity in effects is an indicator of
a vague, poorly specified theory.

Large-scale attempts of direct replications, using identical protocols,
such as the Many Labs studies or Registered Replication Reports [e.g.
@ManyLabs1; @RRRHart], for the first time allow researchers to estimate
heterogeneity undistorted by selection processes like publication bias.
In re-analyses of these studies, Olsson-Collentine et al. [-@olsson2020]
identify a positive correlation between a phenomenon’s effect size and
its heterogeneity. Similarly, van Erp et al. [-@vanErp2017] and Stanley
et al. [-@stanley2018] estimate strong degrees of heterogeneity across
conceptual replications in Psychology. In a separate re-analysis of
large scale direct replications, Renkewitz et al. (in preparation)
identify substantial heterogeneity in almost all projects where a
non-zero effect could be identified. This aligns with the correlation
found by Olsson-Collentine et al. [-@olsson2020].

One source for this heterogeneity, which is not a typical moderator tied
to a specific theory, might be impairments in score reliability which
vary in their extent across replications. \[This is an argument put
forward repeatedly by original researchers and replicators alike (do we
actually know of any?).\] Hunter and Schmidt [-@hunter2004] and Wiernik
and Dahlke [-@wiernik2020], discussing the role of score reliability for
standardized effect sizes, claim that differences in score reliability
should inflate, and therefore increase, heterogeneity. This implies that
standard corrections for imperfect score reliability, such as
attenuation correction procedures, should remove that inflation,
reducing heterogeneity in corrected effect sizes. Over the following
pages, we assess whether score reliability can in fact be made
responsible for heterogeneity identified across phenomena. We find no
empirical evidence for the argument that correcting for differences in
score reliability reduces effect size heterogeneity. Instead, to our
great surprise, we find that heterogeneity might as well be larger after
the correction procedure. Therefore, theoretical analyses regarding the
relationship between score reliability and heterogeneity need to be
false. In the second half of this text we make use of analytical
arguments to identify the necessary conditions under which an
attenuation correction inevitably leads to an increase in effect size
heterogeneity.

@@

Here, MD refers to the mean difference between two groups of interest,
while $\sigma_X$ is the total pooled standard deviation. Since $d$
constitutes an observed effect size it is affected by sampling error. In
terms of heterogeneity, we are not interested in the variance of $d$,
but the variance of the true, underlying ES $\delta$. In a
random-effects meta-analytic model, we can use the standard error and an
estimate of heterogeneity to understand how far off the individual
observed $d$ is from the true $\delta$ in a weighting procedure
[@borenstein2009] The standard error of Cohen’s $d$ is computed as
defined in @eq-SEd.

As is widely known, score reliability affects ES [e.g. @hunter2004;
@wiernik2020]. In the context of classical test theory (CTT) score
reliability $\rho_{XX’}$ is defined as the ratio of true to total score
variance [@lord2008], as defined in @eq-rel.

@@

If information concerning score reliability is available, it is possible
to correct the individual ES for its unreliability. This process is also
known as attenuation correction and, while not without its criticisms
[@muchinsky1996; @schmidt1996; @schennach2016], is a widespread practice
in Psychology (reference). @eq-dc describes a simple attenuation
correction procedure.

@@

$\hat{\sigma}_T$ can be estimated by simply rearranging the terms from
@eq-rel. Thereby, an estimate of ES is generated, under the premise that
score reliability would have been perfect ($\rho_{xx'} = 1$). As the
corrected estimate of Cohen’s $d$ is larger and requires an additional
unknown parameter, an estimate of score reliability, the larger
uncertainty in this parameter should be reflected in a larger standard
error. The standard error for $d_c$ is defined in @eq-SEdc.

@@

-   To what extent can heterogeneity in standardized effect sizes be
    explained by means of an attenuation correction?

@@

We have collected a large number of openly available multi-site direct
replications on psychological phenomena [@dripht2024]. The data-sets of
about 50 phenomena, largely stemming from the efforts of the Many Labs
studies, Registered Replication Reports or the Psychological Science
Accelerator, are made openly available in a standardised format at
\[<https://osf.io/7p8rk/>\], the DRIPHT repository.

@@

The ManyLabs projects were collaborative efforts to replicate several
psychological phenomena across different research sites, employing
identical protocols. From five published projects, the data for the
first three and the fifth Many Labs projects [@ManyLabs1; @ManyLabs2;
@ManyLabs3; @ManyLabs5] was harmonized and publicly stored in a
standardized format when the DRIPHT repository was set up [@dripht2024].
The Registered Replication Reports are similar collaborative efforts,
with the sole distinction that for each report a single phenomenon was
replicated several times. The Registered Reports 3-10 were added to the
DRIPHT repository, as they employed experimental designs with at least
two groups [@RRRHart; @RRRSripada; @RRRFinkel; @RRRStrack; @RRRRand;
@RRRDijksterhuis; @RRRSrull; @RRRMazar]. Lastly, some projects from the
Psychological Science Accelerator were added to the DRIPHT repository,
namely PSA006, PSACR001, PSACR002 and PSACR003 [@PSA006; @PSACR001;
@PSACR002; @PSACR003]. In contrast to Many Labs or Registered
Replication Projects, these collaborative efforts, while also
distributed across the globe, do not formally attempt to perform direct
replications, but focus on original research or conceptual replications.
However, as each project used the same protocol and participants are
distributed across different countries, a data structure similar to that
of a multi-site replication study emerges. What makes all these
collaborative efforts valuable for this manuscript is the fact that all
phenomena collected in the DRIPHT-repository are making use of two-group
designs and are therefore easily assessed using standardised mean
differences.

While replication data on a decent number of psychological phenomena is
available in the DRIPHT-repository, to demonstrate how incorporating
score reliability affects heterogeneity in ES, selected phenomena need
to fulfil three conditions: (1) As we focus on the use of ES $d$, the
phenomenon needs to be assessed using a group design consisting of at
least two groups. (2) Score reliability estimates can be derived, this
implies that the phenomenon needs to be measured using either several
indicators forming a single scale or by repeatedly measuring across
several timepoints. (3) Score reliability can only attenuate effect
sizes that are not zero in the first place. Therefore, we focus on
phenomena where meta-analytic mean ES can be statistically distinguished
from zero. While the first criterion is fulfilled by all phenomena found
in the DRIPHT repository, the second and third criteria still need
evaluation.

@@

From the 21 phenomena that employed more than a single indicator, 19
made use of Likert-style items, where respondents would indicate their
agreement to some kind of statement. The scale-length varied from 3 to
10-points. The items measuring the remaining 2 phenomena were coded
dichotomously, as responses were right or wrong. While the majority of
designs make use of some form of control-treatment manipulation,
randomly assigning participants to a condition (19 phenomena), some
phenomena refer to the effect of pre-existing differences, e.g.
biological sex (2 phenomena). Subsequently, those phenomena need to be
identified where the meta-analytic mean effect size is statistically
significant different from zero. This will be done as a preliminary step
in this manuscript’s analysis procedure. Detailed information concerning
the different phenomena and how they were measured can be found at
\[osf-link\], while a brief summary can be found in Table 1.

@@

@@

For each replication of a phenomenon, Cohen’s $d$ is computed as an
estimate of raw ES, according to @eq-d, including its standard error
(@eq-SEd). Subsequently, a random-effects meta-analysis is performed
using metafor version 4.8-0 [@metafor]. To estimate meta-analytic mean
ES and heterogeneity we make use of the REML-estimator, as that
estimator tends to perform best across a range of conditions
[@hoenekopp2022]. To identify which phenomena pass criterion (3), using
a Wald-type significance test we assess for which phenomenon the
meta-analytic mean ES is significantly different from zero. For this
hypothesis test, as all other hypothesis tests in this manuscript, we
employ a one-tailed significance level of .05.

@@

[Equations @eq-d], [-@eq-SEd], [-@eq-dc], and [-@eq-SEdc] were used to
generate estimates of standardized ES, both uncorrected and corrected,
with their corresponding standard errors. Generally, this leads to
larger (absolute) ES and larger standard errors after the correction. As
an example from the 21 selected phenomena, the results from this
procedure on Nosek’s Explicit Art sex differences phenomenon are
displayed in the forest plot in @fig-forest. Here, grey dots represent
the uncorrected ES for each sample, while black dots represent the
corrected ES in each sample. The bars surrounding these dots show the
respective 95%-Cis.

```{r, echo = FALSE, fig.height = 6}
#| label: fig-forest 
#| #| fig-cap: Forest Plot Nosek_Explicit_Art 
forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
```

@@

@tbl-meanES demonstrates that what we observed in @fig-forest holds
across all 21 phenomena. All meta-analytic effects are larger in their
absolute size after applying an attenuation correction procedure.
Similarly, the uncertainty quantified in the estimate's standard error
is larger. Additionally, @tbl-meanES highlights which phenomena pass
criterion (3)—the effect size must be statistically significantly
distinguishable from zero. 12 phenomena pass this criterion, highlighted
in boldface. It may be remarked that the attenuation correction
procedure, in this data, appears to have no influence on whether a
phenomenon passes the criterion. The differences in p-value are rather
small and lead to no difference in conclusion, regarding a significance
level

```{r, echo = FALSE}
#| label: tbl-meanES
#| tbl-cap: Meta-analytic average  ES
knitr::kable(x = read.csv(here("Tables/Mean_ES.csv")),
             col.names = c("MASC", 
                           "$\\hat{\\mu}_{\\delta_0}$ (SE)", 
                           "p", 
                           "$\\hat{\\mu}_{\\delta}$ (SE)", 
                           "p"),
             escape = FALSE)
```

@@

@tbl-tauES contains the results from performing an RG-MA on the
Cronbach's Alpha coefficients of the scales used to measure the 12
remaining phenomena. Additionally, it contains estimates of
heterogeneity for both uncorrected ES ($\delta_0$) and corrected ES
($\delta$), which are central to assess the research questions posed in
section *Attenuation Correction*. Most importantly, @tbl-tauES
demonstrates that for all but one of the phenomena, the attenuation
correction procedure barely has an impact on the extent of ES
heterogeneity. However, while the heterogeneity for the
Srull_Behaviour_Hostility phenomenon seems reduced by about a third of
its initial value, the ES heterogeneity for this phenomenon is rather
small to begin with. Using a significance test, we can not distinguish
this value confidently from zero. This implies that for the majority of
phenomena, differences in score reliability can not be made responsible
for heterogeneity in standardized effect sizes.

@@

knitr::kable(tab_both,
             col.names = c("MASC", 
                           "$\\hat{\\mu}_{\\rho_{XX'}}$ (95%-CI)", 
                           "$\\hat{\\tau}_{\\rho_{XX'}}$", 
                           "QE (df)", "p", 
                           "$\\hat{\\tau}_{\\delta_0}$", 
                           "QE (df)", "p",
                           "$\\hat{\\tau}_{\\delta}$", 
                           "QE (df)", "p"),
             escape = FALSE)
# 
# tab_one <- data.frame(tab_muES,
#                       tab_hetrel[,c("mu_alpha_str", "tau_alpha", "QE_str", "QEp")])
```

@@

Secondly, in @tbl-tauES we find that for ten out of twelve phenomena, ES
heterogeneity grows as a result from the attenuation correction
procedure. Even though these differences tend to be small, it is clear
that in all those cases, not only did accounting for score reliability
not explain any ES heterogeneity, we find even more than initially
expected. This is incompatible with claims made in W&D and H&S, the
attenuation correction increased ES heterogeneity. Concerning the
remaining two phenomena, for the Shnabel_Willingness_Reconcile_RPP, the
reduction in heterogeneity is miniscule and the estimate of
heterogeneity is not statistically significant in the first place. Most
likely, the difference in ES heterogeneity estimates can be explained by
issues of estimation, as only eight replications are available.
Similarly, only for the Srull_Behaviour_Hostility phenomenon, we find a
substantial reduction in heterogeneity due to the attenuation correction
procedure. The estimates of ES heterogeneity are not statistically
significant, however, which hampers our ability to meaningfully
interpret and compare these estimates. For the majority of phenomena it
appears that differences in score reliability do not inflate
heterogeneity in uncorrected effect sizes. Instead these differences
mask ES heterogeneity, that can be unveiled by accounting for score
reliability.

@@

In the re-analysis of archival data in the previous section we
demonstrated that, at least for the multi-site direct replication and
PSA-projects, that score reliability does not hold any noteworthy value
as a potential moderator explaining ES heterogeneity. Additionally, we
demonstrated that claims formulated in W&D and H&S regarding how score
reliability increases ES heterogeneity is false. We found a large number
of phenomena where, contrary to popular claims, correcting for score
reliability increases ES heterogeneity, instead of decreasing it.

This implies that there is an error in the reasoning or intuition behind
claims made in H&S and W&D, as our observations can not be explained by
the current literature. Over the following paragraphs, we will briefly
explore the intuition that might have led to the previous claims.
Subsequently, we will use analytical arguments to clarify how score
reliability actually affects ES heterogeneity. Therein, we also identify
potential conditions under which an attenuation correction can actually
reduce ES heterogeneity - and under which it is more likely to increase
it.

@@

In the re-analysis of archival data, we have demonstrated that this
intuition does not hold for the relationship of ES heterogeneity and
score reliability. Instead, we will propose an alternative way that
helps us discern, under which conditions we can expect an attenuation
correction procedure to lead to an increase in ES heterogeneity.

@@

Realising that an ES, as defined in @eq-d, is actually a ratio, allows
for a different analytical description of how differences in score
reliability can affect ES heterogeneity. If the phenomenon is
heterogeneous and score reliability varies across replications, it seems
sensible to assume that both numerator and denominator (MD and
$\sigma_X$) from @eq-d vary across replications. In that case, the ES
can be described as a random variable stemming from a ratio
distribution. A ratio distribution is a probability distribution
constructed by dividing one random variable by a second random variable:
$Z = \frac{X}{Y}$. First order taylor approximation may be used to
generate estimates of mean and variance of the ratio variable
\[irrespective of distribution?\] [@vanKempen1999]. Here we use the
Taylor-estimator of the ratio’s variance to demonstrate how differences
in score reliability may affect heterogeneity in ES.

@@

To do so, we introduce two additional metrics.

@@

Estimation of $R_1$ and $R_2$ requires estimates of meta-analytic mean
and heterogeneity of both the total score variance $\sigma^2_X$ and the
true score variance $\sigma^2_T$. Elsewhere, we proposed the method
Boot-Err (ref.) as an alternative method to test for and estimate the
differences in error score variance $\sigma^2_E$. Among other things, in
Beinhauer et al. [-@beinhauer2025] we deliver detailed instructions on
how estimates for mean and heterogeneity of total score variance, and
the score variance components can be derived. This means that Boot-Err
can be used to generate all estimates required for computation of the
metrics $R_1$ and $R_2$.

@@

For the two exceptions, those phenomena where the attenuation correction
procedure actually led to an decrease in ES heterogeneity
(Srull_Behavior_Hostility and Shnabel_Willingness_Reconcile_RPP), the
significance test for the ES heterogeneity was not significant, both
before and after the correction. The results regarding $R_2$ indicate
that the majority of score variance heterogeneity was due to differences
in true score variance $\sigma^2_T$. Since either the heterogeneity was
so small or the sample size so low that the heterogeneity could not be
statistically distinguished from zero, we should be cautious in the
interpretation of a reduction in that estimate. Most likely, the
reduction in heterogeneity is better explained by imprecise estimation
than an actual reduction of heterogeneity.

>>>>

---
title: "Manuscript"
bibliography: ReLiability_references.bib
csl: styles/apa-6th-edition.csl
format: pdf
editor: 
  markdown: 
    wrap: 72
---

@@

# relevant libraries required for this script
packages <- c("ggplot2", "here", "metafor", "gridExtra", "knitr", "magrittr", "dplyr", "gt")

@@

Lukas J. Beinhauer *University of Erfurt* ORCID:
https://orcid.org/0000-0001-9841-3089

Jens H. Fünderich *Ludwig Maximilian University of Munich* ORCID:
https://orcid.org/0000-0002-7185-9248

Frank Renkewitz *University of Erfurt* ORCID:
https://orcid.org/0000-0001-8072-6802

@@

Similarly, it has been argued that heterogeneity in effect sizes is an
indicator of the theory’s “completeness” surrounding the phenomenon. For
example, Linden and Hönekopp argue that “low (as opposed to high)
heterogeneity reflects a more advanced understanding of the subject
matter being studied” [-@LindenHönekopp, p. 2]. Similarly, von Hippel
and Schuetze [-@vonHippel2023] argue that heterogeneity in effects is an
indicator of a vague, poorly specified theory.

Large-scale attempts of direct replications, using identical protocols,
such as the Many Labs studies or Registered Replication Reports
[@ManyLabs1; e.g. @RRRHart], for the first time allow researchers to
estimate heterogeneity undistorted by selection processes like
publication bias. In re-analyses of these studies, Olsson-Collentine et
al. [-@olsson2020] identify a positive correlation between a
phenomenon’s effect size and its heterogeneity. Both van Erp et al.
[-@vanErp2017] and Stanley et al. [-@stanley2018] estimate strong
degrees of heterogeneity across conceptual replications in Psychology.
In a separate re-analysis of large scale direct replications, Renkewitz
et al. (in preparation) identify substantial heterogeneity in almost all
projects where a non-zero effect could be identified. This aligns with
the correlation found by Olsson-Collentine et al. [-@olsson2020].

One source for this heterogeneity, which is not a typical moderator tied
to a specific theory, might be impairments in score reliability which
vary in their extent across replications. \[This is an argument put
forward repeatedly by original researchers and replicators alike, such
as Olsson-Collentine et al. [-@olsson2020] (do we know any more?).\]
Hunter and Schmidt [-@hunter2004] and Wiernik and Dahlke
[-@wiernik2020], discussing the role of score reliability for
standardized effect sizes, claim that differences in score reliability
should inflate, and therefore increase, heterogeneity. This implies that
standard corrections for imperfect score reliability, such as
attenuation correction procedures, should remove that inflation,
reducing heterogeneity in corrected effect sizes.

Over the following pages, we assess whether score reliability can in
fact be made responsible for heterogeneity identified across several
psychological phenomena. However, we find no empirical evidence for the
argument that correcting for differences in score reliability reduces
effect size heterogeneity. Instead, to our great surprise, we find that
heterogeneity might as well be larger after the correction procedure.
Therefore, previous analyses and assumptions regarding the relationship
between score reliability and heterogeneity need to be false. In the
second half of this text we make use of analytical arguments to identify
the necessary conditions under which an attenuation correction
inevitably leads to an increase in effect size heterogeneity \[Do we
need any "Schleifen" in this paragraph - to reduce informational load /
density\].

@@

Here, MD refers to the mean difference between two groups of interest,
while $\sigma_X$ is the pooled standard deviation. As is widely known,
score reliability affects ES [e.g. @hunter2004; @wiernik2020]. In the
context of classical test theory (CTT) score reliability $\rho_{XX’}$ is
defined as the ratio of true to total score variance [@lord2008], as
defined in @eq-rel.

@@

If information concerning score reliability is available, it is possible
to correct the individual ES for its unreliability. This process is also
known as attenuation correction and, while not without its criticisms
[@muchinsky1996; @schmidt1996; @schennach2016], is a widespread practice
in Psychology (reference, alt.: repeatedly recommended + references).
@eq-dc describes a simple attenuation correction procedure.

@@

$\hat{\sigma}_T$ can be estimated by simply rearranging the terms from
@eq-rel. Thereby, an estimate of ES is generated, under the premise that
score reliability would have been perfect ($\rho_{xx'} = 1$).

@@

-   To what extent can heterogeneity in standardized effect sizes be
    reduced by means of an attenuation correction?

@@

We have collected a large number of openly available multi-site direct
replications on psychological phenomena [@dripht2024]. The data-sets of
about 55 phenomena, largely stemming from the efforts of the Many Labs
studies, Registered Replication Reports or the Psychological Science
Accelerator, are made openly available in a standardised format at
\[<https://osf.io/7p8rk/>\], the DRIPHT repository. A documentation of
the standardised format, including a web-application to easily explore
the public data-sets, can be found in Fünderich et al.
[-@fuenderich_metapipex2024].

@@

The ManyLabs projects were collaborative efforts to replicate several
psychological phenomena across different research sites, employing
identical protocols. From five published projects, the data for the
first three and the fifth Many Labs projects [@ManyLabs1; @ManyLabs2;
@ManyLabs3; @ManyLabs5] were harmonized and publicly stored in a
standardized format when the DRIPHT repository was set up [@dripht2024].
The Registered Replication Reports are similar collaborative efforts,
with the sole distinction that for each report a single phenomenon was
replicated several times. The Registered Reports 3-10 were added to the
DRIPHT repository, as they employed experimental designs with at least
two groups [@RRRHart; @RRRSripada; @RRRFinkel; @RRRStrack; @RRRRand;
@RRRDijksterhuis; @RRRSrull; @RRRMazar]. Lastly, some projects from the
Psychological Science Accelerator were added to the DRIPHT repository,
namely PSA006, PSACR001, PSACR002 and PSACR003 [@PSA006; @PSACR001;
@PSACR002; @PSACR003]. In contrast to Many Labs or Registered
Replication Projects, these collaborative efforts, while also
distributed across the globe, do not formally attempt to perform direct
replications. Instead, these projects focus on original research or
conceptual research done collaboratively in laboratories across the
globe. However, as each project used the same protocol and participants
are distributed across different countries, a data structure similar to
that of a multi-site replication study emerges. What makes all these
collaborative efforts valuable for this manuscript is the fact that all
phenomena collected in the DRIPHT-repository are making use of two-group
designs and are therefore easily assessed using standardised mean
differences.

While replication data on a decent number of psychological phenomena is
available in the DRIPHT-repository, selected phenomena need to fulfil
three conditions to demonstrate how incorporating score reliability
affects heterogeneity in ES: (1) As we focus on the use of ES $d$, the
phenomenon needs to be assessed using a group design consisting of at
least two groups. (2) Score reliability estimates can be derived, this
implies that the phenomenon needs to be measured using either several
indicators forming a single scale or by repeatedly measuring across
several timepoints. (3) Score reliability can only attenuate effect
sizes that are not zero in the first place. Therefore, we focus on
phenomena where meta-analytic mean ES can be statistically distinguished
from zero. This aligns with the observation made in (Renkewitz,
Fünderich & Beinhauer, in preparation) and Olsson-Collentine et al.
[-@olsson2020]. As null-effects displayed little to no heterogeneity,
there is no variability that might be reduced by incorporating
differences in score reliability. While the first criterion is fulfilled
by all phenomena found in the DRIPHT repository, the second and third
criteria still need evaluation.

@@

From the 21 phenomena that employed more than a single indicator, 19
made use of Likert-style items, where respondents would indicate their
agreement to some kind of statement. The scale-length varied from 3 to
10-points. The items measuring the remaining 2 phenomena were coded
dichotomously, as responses were right or wrong. While the majority of
designs make use of some form of control-treatment manipulation,
randomly assigning participants to a condition (19 phenomena), some
phenomena refer to the effect of pre-existing differences, e.g.
biological sex (2 phenomena). Subsequently, those phenomena need to be
identified where the meta-analytic mean effect size is statistically
significant different from zero. This will be done as a preliminary step
in this manuscript’s analysis procedure. Detailed information concerning
the different phenomena and how they were measured can be found at
\[osf-link\], while a brief summary can be found in @tbl-data.

@@

In order to derive estimates of meta-analytic heterogeneity, we make use
of a random-effects meta-analysis. Broken down, a random-effects
meta-analysis essentially pools the sampling uncertainty of the
different estimates that are to be aggregated. By removing this pooled
estimate from the observed variability in the effect size estimates, we
can generate an estimate of its heterogeneity - how strongly the effect
sizes vary beyond their pooled sampling uncertainty.

For the standardized mean difference $d$, we can compute the standard
error as an estimate of its sampling uncertainty. The standard error of
Cohen’s $d$ is computed as defined in @eq-SEd.

As the corrected estimate of Cohen’s $d$ is larger and requires an
additional unknown parameter, an estimate of score reliability, the
larger uncertainty in this parameter should be reflected in a larger
standard error. The standard error for $d_c$ is defined in @eq-SEdc.

@@

For each replication of a phenomenon, Cohen’s $d$ is computed as an
estimate of raw ES, according to @eq-d, including its standard error
(@eq-SEd). Subsequently, a random-effects meta-analysis is performed
using metafor version 4.8-0 [@metafor]. To estimate meta-analytic mean
ES and heterogeneity we make use of the REML-estimator, as that
estimator tends to perform best across a range of conditions
[@hoenekopp2022]. To identify which phenomena pass criterion (3) we use
a Wald-type significance test we assess for which phenomenon the
meta-analytic mean ES is significantly different from zero. For this
hypothesis test, as all other hypothesis tests in this manuscript, we
employ a one-tailed significance level of .05.

@@

[Equations @eq-d], [-@eq-SEd], [-@eq-dc], and [-@eq-SEdc] were used to
generate estimates of standardized ES, both uncorrected and corrected,
with their corresponding standard errors. Generally, this leads to
larger (absolute) ES and larger standard errors after the correction. As
an example from the 21 selected phenomena, the results from this
procedure on replications on Nosek’s study on sex differences in
preferences for the subject of Arts is displayed in the forest plot in
@fig-forest. Here, grey dots represent the uncorrected ES for each
sample, while black dots represent the corrected ES in each sample. The
bars surrounding these dots show the respective 95%-Cis.

```{r, echo = FALSE, fig.height = 6}
#| label: fig-forest 
#| fig-cap: Forest Plot ML1 - Nosek (Art)
forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
```

@@

@tbl-meanES demonstrates that what we observed in @fig-forest holds
across all 21 phenomena. All meta-analytic effects are larger in their
absolute size after applying an attenuation correction procedure.
Similarly, the uncertainty quantified in the estimate's standard error
is larger. Additionally, @tbl-meanES highlights which phenomena pass
criterion (3)—the effect size must be statistically significantly
distinguishable from zero at $p < .05$. 12 phenomena pass this
criterion. It may be remarked that the attenuation correction procedure,
in this data, appears to have no influence on whether a phenomenon
passes the criterion. The differences in p-value are rather small and
lead to no difference in conclusion, regarding a significance level of
5%.

```{r, echo = FALSE}
#| label: tbl-meanES
#| tbl-cap: Meta-analytic average  ES

gt::gt(read.csv(here("Tables/Mean_ES.csv")) %>% 
         arrange(MASC)) %>% 
  gt::cols_label(MASC = "MASC",
                 mu_str_raw = gt::md("$\\hat{\\mu}_{\\delta_0}$ (SE)"),
                 pval_raw = "p",
                 mu_str_cor = gt::md("$\\hat{\\mu}_{\\delta}$ (SE)"),
                 pval_cor = "p") %>% 
  gt::tab_options(table.width = gt::pct(100),
                  column_labels.font.weight = "bold") # %>% 
  # gt::cols_width(pval_raw ~ gt::px(40),
  #                pval_cor ~ gt::px(40))
```

@@

@tbl-tauES contains the results from performing an RG-MA on the
Cronbach's Alpha coefficients of the scales used to measure the 12
remaining phenomena. Additionally, it contains estimates of
heterogeneity for both uncorrected ES ($\delta_0$) and corrected ES
($\delta$), which are central to assess the research questions posed in
section *Attenuation Correction*. Most importantly, @tbl-tauES
demonstrates that for all but one of the phenomena, the attenuation
correction procedure barely has an impact on the extent of ES
heterogeneity. However, while the heterogeneity for the phenomenon from
RRR9 (Behaviour) seems reduced by about a third of its initial value,
the ES heterogeneity for this phenomenon is rather small to begin with.
Using a significance test, we can not distinguish this value confidently
from zero. This implies that for the majority of phenomena, differences
in score reliability can not be made responsible for heterogeneity in
standardized effect sizes.

@@

# knitr::kable(tab_both,
#              col.names = c("MASC", 
#                            "$\\hat{\\mu}_{\\rho_{XX'}}$ (95%-CI)", 
#                            "$\\hat{\\tau}_{\\rho_{XX'}}$", 
#                            "QE (df)", "p", 
#                            "$\\hat{\\tau}_{\\delta_0}$", 
#                            "QE (df)", "p",
#                            "$\\hat{\\tau}_{\\delta}$", 
#                            "QE (df)", "p"),
#              escape = FALSE)

gt::gt(tab_both  %>% 
  arrange(tab_hetES.MASC)) %>% 
  gt::cols_label(tab_hetES.MASC = "MASC",
                 mu_alpha_str = gt::md("$\\hat{\\mu}_{\\rho_{XX'}}$"),
                 tau_alpha = gt::md("$\\hat{\\tau}_{\\rho_{XX'}}$"),
                 QE_str = "QE (df)",
                 QEp = "p",
                 tau_raw = gt::md("$\\hat{\\tau}_{\\delta_0}$"),
                 QE_raw_str = "QE (df)",
                 QEp_raw = "p",
                 tau_raw = gt::md("$\\hat{\\tau}_{\\delta}$"),
                 QE_raw_str = "QE (df)",
                 QEp_raw = "p") %>% 
  gt::tab_options(table.width = gt::pct(100),
                  column_labels.font.weight = "bold") #%>% 
  # gt::cols_width(pval_raw ~ gt::px(40),
  #                pval_cor ~ gt::px(40))

# tab_one <- data.frame(tab_muES,
#                       tab_hetrel[,c("mu_alpha_str", "tau_alpha", "QE_str", "QEp")])
```

@@

Secondly, in @tbl-tauES we find that for ten out of twelve phenomena, ES
heterogeneity grows as a result from the attenuation correction
procedure. Even though these differences tend to be small, it is clear
that in all those cases, not only did accounting for score reliability
not explain any ES heterogeneity, we find even more than initially
expected. This is incompatible with claims made in W&D and H&S, the
attenuation correction increased ES heterogeneity. Concerning the
remaining two phenomena, for the phenomenon described in ML5 - Shnabel
(RPP), the reduction in heterogeneity is miniscule and the estimate of
heterogeneity is not statistically significant in the first place. Most
likely, the difference in ES heterogeneity estimates can be explained by
issues of estimation, as only eight replications are available.
Similarly, only for the phenomenon described in RRR9 (Behaviour), we
find a seemingly substantial reduction in heterogeneity due to the
attenuation correction procedure. The estimates of ES heterogeneity are
not statistically significant, however, which hampers our ability to
meaningfully interpret and compare these estimates. For the majority of
phenomena it appears that differences in score reliability do not
inflate heterogeneity in uncorrected effect sizes. Instead these
differences mask ES heterogeneity, that can be unveiled by accounting
for score reliability.

@@

In the re-analysis of archival data in the previous section we
demonstrated that, at least for the multi-site direct replication and
PSA-projects, score reliability does not hold any noteworthy value as a
potential moderator explaining ES heterogeneity. Additionally, we
demonstrated that claims formulated in W&D and H&S regarding how score
reliability increases ES heterogeneity are false. We found a large
number of phenomena where, contrary to popular claims, correcting for
score reliability increases ES heterogeneity, instead of decreasing it.

This implies that there is an error in the reasoning or intuition behind
claims made in H&S and W&D, as our observations can not be explained by
the current literature. Over the following paragraphs, we will briefly
explore the intuition that might have led to the previous claims.
Subsequently, we will use analytical arguments to clarify how score
reliability actually affects ES heterogeneity. Therein, we also identify
potential conditions under which an attenuation correction will actually
increase ES heterogeneity - and under which it is more likely to
decrease it.

@@

In the re-analysis of archival data, we have demonstrated that this
intuition does not hold for the relationship of ES heterogeneity and
score reliability. Instead, we propose an alternative way that helps us
discern, under which conditions we can expect an attenuation correction
procedure to lead to an increase in ES heterogeneity.

@@

Realising that an ES, as defined in @eq-d, is actually a ratio, allows
for a different analytical description of how differences in score
reliability can affect ES heterogeneity. If the phenomenon is
heterogeneous and score reliability varies across replications, it seems
sensible to assume that both numerator and denominator (MD and
$\sigma_X$) from @eq-d vary across replications. In that case, the ES
can be described as a random variable stemming from a ratio
distribution. A ratio distribution is a probability distribution
constructed by dividing one random variable by a second random variable:
$Z = \frac{X}{Y}$. First order taylor approximation may be used to
generate estimates of mean and variance of the ratio variable
[@vanKempen1999]. Taylor approximations typically require no specific
assumptions regarding the distribution form, but that the variable (here
a ratio) is a) differentiable, b) has existing moments and that c) its
higher-order moments are negligible (e.g. limited variability and
monotonicity) (still need a good reference here). Here we use the
Taylor-estimator of the ratio’s variance to demonstrate how differences
in score reliability may affect heterogeneity in ES.

@@

To do so, we had to introduce two additional metrics.

@@

Estimation of $R_1$ and $R_2$ requires estimates of meta-analytic mean
and heterogeneity of both the total score variance $\sigma^2_X$ and the
true score variance $\sigma^2_T$. Elsewhere, we proposed the method
Boot-Err [@beinhauer2025] as an alternative method to test for and
estimate the differences in error score variance $\sigma^2_E$. Among
other things, in Beinhauer et al. [-@beinhauer2025] we deliver detailed
instructions on how estimates for mean and heterogeneity of total score
variance, and the score variance components can be derived. This means
that Boot-Err can be used to generate all estimates required for
computation of the metrics $R_1$ and $R_2$.

@@

For the two exceptions, those phenomena where the attenuation correction
procedure actually led to a decrease in ES heterogeneity (RRR9
(Behaviour) and ML5 - Shnabel (RPP)), the significance test for the ES
heterogeneity was not significant, both before and after the correction.
The results regarding $R_2$ indicate that the majority of score variance
heterogeneity was due to differences in true score variance
$\sigma^2_T$. Since either the heterogeneity was so small or the sample
size so low that the heterogeneity could not be statistically
distinguished from zero, we should be cautious in the interpretation of
a reduction in that estimate. Most likely, the reduction in
heterogeneity is better explained by imprecise estimation than an actual
reduction of heterogeneity.

Lastly, for two out of the 12 phenomena, the one described in ML5 -
Albarracin and the other described in ML5 - Shnabel (Rev), we find
$\tau_{\sigma^2_X}$ of 0. Therefore, the $R_2$ metric could not be
computed, as the denominator would have been zero. However, we also did
not identify statistically significant ES heterogeneity in the first
place. The phenomena were assessed with a particularly low number of
replications (9 and 8 respectively). Estimation precision for
heterogeneity in variance components, even more so than for ES, largely
depends on the sample size. Most likely, the low sample size led to
implausible estimates in both true and total score variance
heterogeneity and subsequently to an implausible estimate of $R_2$.
Meta-analytic estimates, especially those concerning score variance
components such as $R_2$, therefore should be interpreted with caution,
if the sample size was particularly low.
